---
title: "R Notebook Example"
output:
  pdf_document: default
  html_document: default
date: "2023-11-24"
---

To insert rchunk - press C button below name

Alternatively you  can just use ```escape! ```  You can also set up figures etc.

```{r first_chunk, include= TRUE, echo = TRUE, message = FALSE, warning=TRUE}
library(tidyverse)
gapminder_frame = read_csv('data/gapminder_data.csv')
a = 4
print (a)
print(gapminder_frame)
```
```{r load_data, include= TRUE, echo = FALSE, message = FALSE, warning=FALSE}
library(tidyverse)
gapminder_frame = read_csv('data/gapminder_data.csv')
```


Create Data table from Europe in 2007, and sort with largest lfie expectacny at top
```{r filter_data}
euro_table_2007 = gapminder_frame %>% filter (continent == "Europe" & year == 2007) %>% 
  select(-continent, -year) %>% 
  arrange(desc(lifeExp)) %>% 
  rename (Country = country, "Population Size" = pop, "Life Expectancie" = lifeExp, "GDP (per capita)" = gdpPercap)
```


Now create table fucntion 
```{r table_maker_kbl}
library(kableExtra)
euro_table_2007 %>% kbl(caption = "Europe table gdp and life expectancy") %>% kable_styling(bootstrap_options = c("striped", "hover"))
```

bootstrap_options Options Include
Striped (make them striped)
hover(shimmer whe you put mouse ove rit)
condensed (make smaller)
responsive (scrollable on phone)

full_width - ?size of table
Position


```{r paged_maker_kbl}
library(rmarkdown)

euro_table_2007 %>% paged_table()
```

(skip the bit about images/optinals)
Plotting and shit
```{r plot, message = FALSE}
library(ggplot2)
gapminder_euro = gapminder_frame %>% filter (continent == "Europe") %>% mutate(year=as_factor(year))
europlot = ggplot(data=gapminder_euro, aes(x = year, y= lifeExp, label = "EU Violin")) + geom_violin() +
                stat_summary(fun = median, geom = "point")
```
```{r ukplot}
library(ggpubr)
gapminder_gb = gapminder_frame %>% filter (country == "United Kingdom")
gbplot = ggplot (data = gapminder_gb, aes (x=year, y = lifeExp, label ="GB")) + geom_point()
ggarrange(gbplot, europlot, labels = c("A","B"))
```